---
output:
  html_document:
    keep_md: false
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
params:
  set_title: Statistics and Forecast Skill Visualisations
---

---
title: `r params$set_title`
---

```{r setup, include="FALSE"}
#'========================================================================
# Visualise skill metrics
#'========================================================================
#
# by Mark R Payne
# DTU-Aqua, Kgs. Lyngby, Denmark
# http://www.staff.dtu.dk/mpay
#
# Created Wed May  1 10:48:18 2019
#
# Makes a series of generic skill visualisations
#
# This work is subject to a Creative Commons "Attribution" "ShareALike" License.
# You are largely free to do what you like with it, so long as you "attribute"
# me for my contribution. See the fine print at the end for exact details.
#
# To do:
#
# Notes:
#   Make sure when knitting that you have the Knit directory set to "project directory"
#'========================================================================

#Load packages
library(PredEng)
library(tidyverse)
library(sf)
library(here)
#library(scales)

#GGplot config
theme_set(theme_bw(base_size = 14))

#Configure Knitr
knitr::opts_chunk$set(
  echo=FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	results = "markup",
	fig.width = 10,
	fig.height=14/25*10,
  fig.align="center")

#Powerpoint needs 14 x 25cm figures

#knitr::opts_knit$set(root.dir=getwd())
#cat(sprintf("title: %s skill visualisation\n",pcfg@project.name))

```

by Mark R Payne  <br>
DTU-Aqua, Kgs. Lyngby, Denmark  <br>
http://www.staff.dtu.dk/mpay  <br>
```{r results="asis"}
cat(sprintf("Visualisations generated  %s\n",date()))
```


## Configuration

```{r}
#==========================================================================
# Configure
#==========================================================================
pcfg <- PE.load.config()
PE.config.summary(pcfg)
plot(pcfg)

start.near.term <- 2015

#'========================================================================
# Setup ####
#'========================================================================
#Import data
stats.tbl <- PE.db.tbl(pcfg,PE.cfg$db$stats,src=NULL)
mets.tbl <- PE.db.tbl(pcfg,PE.cfg$db$metrics,src=NULL)

# #Plot misc!
# x.lbls <- tibble(months=c(seq(0,12,by=3),seq(24,120,by=24)),t=rescale.fn(months)) %>%
#   mutate(lbl=ifelse(months<=12,sprintf("%g m",months),
#                     sprintf("%g y",months/12)))
# x.minor <- tibble(months=c(0:12,seq(24,120,by=12)),t=rescale.fn(months))


```



```{r results="asis"}
#'========================================================================
# Stat skills ####
#'========================================================================
#Present results by spatial region
sp.names <-
  stats.tbl %>%
  distinct(spName) %>%
  collect() %>%
  pull(spName)

for(this.sp in sp.names) {
  #Show title
  cat(sprintf("\n\n## %s\n",gsub("_"," ",this.sp)))
  
  #Now split it down further into Individual statistics and loop
  stat.names <-
    stats.tbl %>%
    filter(spName==this.sp) %>%
    distinct(statName) %>%
    collect() %>%
    pull(statName)
  

  for(this.stat in stat.names) {
    # Stats ---------------------------------------------------------------------------
    #Loop over resultNames as well
    result.names <- 
      stats.tbl %>%
      filter(spName==this.sp,
             statName==this.stat) %>%
      distinct(resultName) %>%
      collect() %>%
      pull(resultName) %>%
      sort()
    
    # Results -------------------------------------------------------------------------
    for(this.result in result.names) {
      #Show a subtitle
      cat(sprintf("\n\n### %s / %s\n",this.stat,this.result))
      
      #Extract the corresponding reference stat data 
      stat.ref <- 
        stats.tbl %>%
        filter(srcType=="Observations",
               realization=="realmean",
               calibrationMethod=="None",
               spName==this.sp,
               statName==this.stat,
               resultName==this.result) %>% 
        collect() %>%
        mutate(date=ymd(date),
               lbl="Reference",
               highlight=TRUE) %>%
        filter(month(date) %in% pcfg@MOI) %>%
        dplyr::select(-calibrationMethod,-srcType)
        
      #Extract the first predictions
      stat.pred <- 
        stats.tbl %>%
        filter(srcType %in% c("Decadal","NMME","Observations"),
               spName==this.sp,
               statName==this.stat,
               resultName==this.result,
               realization %in% c("realmean","ensmean","grandens"),
               calibrationMethod!="None") %>%
        collect() %>% #Retrieve all
        group_by(srcType,srcName) %>%
        filter(lead==min(lead) | is.na(lead),  #Select the first prediction
               month(date) %in% pcfg@MOI) %>%
        mutate(type.name=paste(srcType,srcName,sep="."),
               date=ymd(date),
               lbl=paste(srcType,srcName,sep="/"),
               highlight=realization %in% c("ensmean","grandens")) %>%
        ungroup()

      #Plot timeseries
      g <- 
        ggplot(mapping=aes(x=date,y=value,colour=lbl,size=highlight))+
        annotate("rect",
                 xmin=as.Date(ISOdate(min(pcfg@comp.years),1,1)),
                 xmax=as.Date(ISOdate(max(pcfg@comp.years)+1,1,1)),
                 ymin=-Inf,ymax=Inf,
                 fill="grey90")+
        geom_line(data=stat.ref)+
        geom_line(data=stat.pred)+
        labs(x="Year",y=this.stat)+
        facet_grid(srcType~calibrationMethod) +
        scale_size_manual(values=c(0.5,2))+
        labs(colour="Forecast system")+
        guides(size="none")
      print(g) 

      # Metrics -------------------------------------------------------------------------
      metric.names <-
        mets.tbl %>%
        filter(spName==this.sp,
               statName==this.stat,
               resultName==this.result,
               metric!="n") %>%   #Don't really need to see this
        distinct(metric) %>%
        collect() %>%
        pull(metric) %>%
        sort() %>%
        factor() %>%
        relevel(ref="pearson.correlation") %>%
        levels()

      #Iterate over skill metrics
      for(this.met in metric.names) {
        #Get data
        all.met.dat <- 
          mets.tbl %>%
          filter(spName==this.sp,
                 statName==this.stat,
                 resultName==this.result,
                 metric==this.met,
                 realization %in% c("realmean","ensmean","grandens") | srcType=="Persistence") %>% 
          collect() %>%
          mutate(type.name=paste(srcType,srcName,sep="/"),
                 highlight=realization %in% c("ensmean","grandens")|srcType=="Persistence") 
        
        met.dat <- 
          all.met.dat %>%
          filter(! srcType %in% c("Persistence","Climatology"))
        
        met.persis.dat <-
          all.met.dat %>%
          filter(srcType %in% c("Persistence","Climatology")) %>%
          dplyr::select(-calibrationMethod)
        
        #Set y range accordingly
        if(this.met %in% c("pearson.correlation")) {
          y.rng <- c(0,1)
        } else if(this.met %in% c("MSSS","crpss")) {
          y.rng <- c(-1,1) 
        } else {
          y.rng <- range(pretty(c(0,met.dat$value)))
        }
        
        #Plot - but only if there is something in the met.dat. For example, it is not 
        #possible to calculate a crps metric for a stat_mean or stat_median -  this would
        #require a bit more developed code
        if(nrow(met.dat)!=0) {
          g <- ggplot(mapping=aes(x=lead/12,y=value,
                                  group=type.name,
                                  colour=type.name,
                                  size=highlight))+
            geom_line(data=met.persis.dat)+
            geom_line(data=met.dat)+
            facet_wrap(~calibrationMethod,scales = "free_y")+
            coord_cartesian(ylim=y.rng)+
            scale_size_manual(values=c(0.25,1))+
            guides(size="none")+
            labs(x="Lead (years)",y=this.met,colour="Forecast System")
          print(g)
        }
        
      }  #Plot  
    } 
  }
  

}


dbDisconnect(stats.tbl)
dbDisconnect(mets.tbl)
```





***
*This work by Mark R Payne is licensed under a  Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported License. 
For details, see http://creativecommons.org/licenses/by-nc-sa/3.0/deed.en_US
Basically, this means that you are free to "share" and "remix" for 
non-commerical purposes as you see fit, so long as you "attribute" me for my
contribution. Derivatives can be distributed under the same or 
similar license.*

*This work comes with ABSOLUTELY NO WARRANTY or support.*

***



