---
title: "Visualise Field skill metrics"
output:
  html_document: 
    df_print: paged
    keep_md: no
    toc: yes
    toc_float: yes
---

```{r setup, include="FALSE"}
#'========================================================================
# Field_skill_metrics
#'========================================================================
#
# by Mark R Payne
# DTU-Aqua, Kgs. Lyngby, Denmark
# http://www.staff.dtu.dk/mpay
#
# Created Wed Mar 24 09:19:59 2021
#
# <Description>
#
# This work is subject to a Creative Commons "Attribution" "ShareALike" License.
# You are largely free to do what you like with it, so long as you "attribute"
# me for my contribution. See the fine print at the end for exact details.
#
# To do:
#
# Notes:
#   * Knitting of notebooks from Rstudio leads to all sorts of messy problems 
#     with the directory. Use here::here() to specify a location relative to
#     the project top directory
#
#'========================================================================

#Configure Knitr
knitr::opts_chunk$set(
  echo=FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	results = "markup",
	fig.width = 10,
	fig.height=14/25*10,
  fig.align="center",
  fig.path=here::here("notebooks","figures","Field_skill_metrics_"))

#Powerpoint needs 14 x 25cm figures

# knitr::opts_knit$set(
#    root.dir="..")

library(PredEng)
library(here)

theme_set(theme_bw(base_size = 14))
```

```{r}
#==========================================================================
# Configure ####
#==========================================================================

```

by Mark R Payne  <br>
DTU-Aqua, Kgs. Lyngby, Denmark  <br>
http://www.staff.dtu.dk/mpay  <br>
```{r results="asis"}
cat(sprintf("Visualisations generated  %s\n",date()))
```

This notebook visualises the skill metrics for statistics that return a field, rather than a single scalar value.

## Configuration

```{r}
#==========================================================================
# Configure
#==========================================================================
pcfg <- PE.load.config()
PE.config.summary(pcfg)
plot(pcfg)

#'========================================================================
# Setup ####
#'========================================================================
#Import data
mets.tbl <- PE.db.tbl(pcfg,PE.cfg$db$metrics.field,src=NULL)

```


```{r results="asis"}
#'========================================================================
# Stat skills ####
#'========================================================================
#Present results by spatial region
sp.names <-
  mets.tbl %>%
  distinct(spName) %>%
  collect() %>%
  pull(spName)

for(this.sp in sp.names) {
  #Show title
  cat(sprintf("\n\n## %s\n",gsub("_"," ",this.sp)))
  
  #Now split it down further into Individual statistics and loop
  stat.names <-
    mets.tbl %>%
    filter(spName==this.sp) %>%
    distinct(statName) %>%
    collect() %>%
    pull(statName)
  

  for(this.stat in stat.names) {
    # Stats ---------------------------------------------------------------------------
    #Loop over resultNames as well
    result.names <- 
      mets.tbl %>%
      filter(spName==this.sp,
             statName==this.stat) %>%
      distinct(resultName) %>%
      collect() %>%
      pull(resultName)
    
    # Results -------------------------------------------------------------------------
    for(this.result in result.names) {
      #Show a subtitle
      cat(sprintf("\n\n### %s / %s\n",this.stat,this.result))
 
      # Metrics -------------------------------------------------------------------------
      metric.names <-
        mets.tbl %>%
        filter(spName==this.sp,
               statName==this.stat,
               resultName==this.result,
               metric!="n") %>%   #Don't really need to see this
        distinct(metric) %>%
        collect() %>%
        filter(!str_ends(metric,".draws")) %>% #Don't take the draws
        pull(metric) 
      
      #Iterate over skill metrics
      for(this.met in metric.names) {
        #Get data
        met.dat <- 
          mets.tbl %>%
          filter(spName==this.sp,
                 statName==this.stat,
                 resultName==this.result,
                 metric==this.met,
                 realization %in% c("realmean","ensmean","grandens")) %>% 
          collect() %>%
          mutate(type.name=paste(srcType,srcName,sep="/"),
                 highlight=realization %in% c("ensmean","grandens"))
        
        #Set y range accordingly
        if(this.met %in% c("pearson.correlation")) {
          y.rng <- c(0,1)
        } else if(this.met %in% c("MSSS","crpss")) {
          y.rng <- c(-1,1) 
        } else {
          y.rng <- range(pretty(c(0,met.dat$value)))
        }
        
        #Plot the first predictions for each srcName, as a way of looking at the initialisation
        g <- 
          met.dat %>%
          #filter(!srcName%in%c("grandens","ensmean")) %>%
          group_by(srcName) %>%
          filter(lead==min(lead)) %>%
          PE.db.unserialize() %>%
          mutate(field=map(field,~ as_tibble(rasterToPoints(.x))),
                 field=map(field,~set_names(.x,nm=c("lon","lat","value")))) %>%
          unnest(field) %>%
          ggplot(aes(lon,lat,fill=value))+
          geom_raster()+
          annotation_map(map_data("world"),fill="black")+
          scale_fill_viridis_c(limits=c(0,0.025))+
          facet_wrap(~type.name)+
          ggtitle(sprintf("First forecast, %s",this.met))+
          coord_quickmap(expand=FALSE)+
          labs(x="",y="")
        print(g)
        


      }  #Plot  
    } 
  }
  

}


dbDisconnect(mets.tbl)
```






***
*This work by Mark R Payne is licensed under a  Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported License. 
For details, see http://creativecommons.org/licenses/by-nc-sa/3.0/deed.en_US
Basically, this means that you are free to "share" and "remix" for 
non-commerical purposes as you see fit, so long as you "attribute" me for my
contribution. Derivatives can be distributed under the same or 
similar license.*

*This work comes with ABSOLUTELY NO WARRANTY or support.*

***



