---
output:
  html_document:
    keep_md: false
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
params:
  set_title: NAO Matching skill
---

---
title: `r params$set_title`
---

```{r setup, include="FALSE"}
#'========================================================================
# Visualise skill metrics
#'========================================================================
#
# by Mark R Payne
# DTU-Aqua, Kgs. Lyngby, Denmark
# http://www.staff.dtu.dk/mpay
#
# Created Wed May  1 10:48:18 2019
#
# Makes a series of generic skill visualisations
#
# This work is subject to a Creative Commons "Attribution" "ShareALike" License.
# You are largely free to do what you like with it, so long as you "attribute"
# me for my contribution. See the fine print at the end for exact details.
#
# To do:
#
# Notes:
#   Make sure when knitting that you have the Knit directory set to "project directory"
#'========================================================================

#Load packages
library(tidyverse)
library(PredEng)
library(sf)
library(here)
#library(scales)

#GGplot config
theme_set(theme_bw(base_size = 14))

#Configure Knitr
knitr::opts_chunk$set(
  echo=FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	results = "markup",
	fig.width = 10,
	fig.height=14/25*10,
  fig.align="center")

#Powerpoint needs 14 x 25cm figures

#knitr::opts_knit$set(root.dir=getwd())
#cat(sprintf("title: %s skill visualisation\n",pcfg@project.name))

```

by Mark R Payne  <br>
DTU-Aqua, Kgs. Lyngby, Denmark  <br>
http://www.staff.dtu.dk/mpay  <br>
```{r results="asis"}
cat(sprintf("Visualisations generated  %s\n",date()))
```


## Configuration

```{r}
#==========================================================================
# Configure
#==========================================================================
pcfg <- PE.load.config()
PE.config.summary(pcfg)

#'========================================================================
# Setup ####
#'========================================================================
#Import data
this.db <- PE.db.connection(pcfg,results.db = TRUE)
stats.tbl <- tbl(this.db,PE.cfg$db$stats)
mets.tbl <- tbl(this.db,PE.cfg$db$metrics)

#Rescale time axis
rescale.fn <- function(x,b) {  #Creates axes that change in linear timescale
  ifelse(x>b[2],
         rescale(x,from=b[2:3],to=c(1,2)),
         rescale(x,from=b[1:2],to=c(0,1)))
}


# #Plot misc!
# x.lbls <- tibble(months=c(seq(0,12,by=3),seq(24,120,by=24)),t=rescale.fn(months)) %>%
#   mutate(lbl=ifelse(months<=12,sprintf("%g m",months),
#                     sprintf("%g y",months/12)))
# x.minor <- tibble(months=c(0:12,seq(24,120,by=12)),t=rescale.fn(months))


```



```{r results="asis"}
#'========================================================================
# Metrics skills ####
#'========================================================================
#Establish which metrics we have NAOmatching calculated for
NAOmatch.methods <-
  mets.tbl %>%
  distinct(calibrationMethod) %>%
  collect() %>%
  filter(grepl("^NAOmatching",calibrationMethod)) %>%
  pull() 

#Import data
NAO.matched <-
  mets.tbl %>%
  filter(calibrationMethod %in% NAOmatch.methods) %>%
  collect() %>%
  separate(calibrationMethod,c("calibration","matching","members"))

#Get unique combinations
uniq <-
  NAO.matched %>%
  distinct(spName,statName,resultName) %>%
  split(.$spName)

for(this.sp in names(uniq)) {
  #Show title
  cat(sprintf("\n\n## %s\n",gsub("_"," ",this.sp)))
  
  #Map of the Spatial region
  #May need to handle global case here as well???
  sp.obj <- 
    pcfg@spatial.polygons %>%
    filter(name==this.sp)
  assert_that(nrow(sp.obj)==1,msg = "Must be one unique spatial object")
  g <- ggplot()+
    geom_sf(data=sp.obj) +
    annotation_map(map_data("world"),fill="black",colour="grey")
  print(g)
  
  #Now split it down further into Individual stats and results and loop
  stat.result <-
    uniq[[this.sp]] %>%
    mutate(stat.res=paste(statName,resultName,sep=" / ")) %>%
    pull(stat.res)

    # Results -------------------------------------------------------------------------
    for(this.statres in stat.result) {
    
      # Metrics -------------------------------------------------------------------------
      metric.names <-
        mets.tbl %>%
        mutate(statres=paste(statName,resultName,sep=" / ")) %>%
        filter(spName==this.sp,
               statres==this.statres,
               metric!="n") %>%   #Don't really need to see this
        distinct(metric) %>%
        collect() %>%
        pull(metric)
      
      #Iterate over skill metrics
      for(this.met in metric.names) {
        #Get data
        all.met.dat <- 
          mets.tbl %>%
          mutate(statres=paste(statName,resultName,sep=" / ")) %>%
          filter(spName==this.sp,
                 statres==this.statres,
                 metric==this.met,
                 realization %in% c("realmean","ensmean") | srcType=="Persistence") %>% 
          collect() %>%
          mutate(type.name=paste(srcType,srcName,sep="."),
                 highlight=realization=="ensmean"|srcType=="Persistence") %>%
          separate(calibrationMethod,c("calibration","matching","members"),fill="right") 

        
        met.dat <- 
          all.met.dat %>%
          filter(srcType!="Persistence",
                 calibration=="NAOmatching")        
        
        met.persis.dat <-
          all.met.dat %>%
          filter(srcType=="Persistence") %>%
          dplyr::select(-calibration,-members,-matching,-type.name)  #So that it is added to all facets
        
        
        #Set y range accordingly
        if(this.met=="pearson.correlation") {
          y.rng <- c(0,1)
        } else {
          y.rng <- range(pretty(c(0,met.dat$value)))
        }
        
        #Plot
        g <- ggplot(mapping=aes(x=lead/12,y=value))+
          geom_line(data=met.persis.dat)+
          geom_line(data=met.dat,mapping=aes(colour=members))+
          facet_grid(type.name~matching,scales = "free_y")+
          coord_cartesian(ylim=y.rng)+
          scale_size_manual(values=c(0.25,1))+
          guides(size="none")+
          labs(x="Lead (years)",y=this.met,colour="Number of members")
        print(g)
        
        g <- ggplot(mapping=aes(x=lead/12,y=value))+
          geom_line(data=met.persis.dat)+
          geom_line(data=met.dat,mapping=aes(colour=matching))+
          facet_grid(type.name~members,scales = "free_y")+
          coord_cartesian(ylim=y.rng)+
          scale_size_manual(values=c(0.25,1))+
          guides(size="none")+
          labs(x="Lead (years)",y=this.met,colour="Matching method")
        print(g)

      }  #Plot  
    } 
  }
  



```





***
*This work by Mark R Payne is licensed under a  Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported License. 
For details, see http://creativecommons.org/licenses/by-nc-sa/3.0/deed.en_US
Basically, this means that you are free to "share" and "remix" for 
non-commerical purposes as you see fit, so long as you "attribute" me for my
contribution. Derivatives can be distributed under the same or 
similar license.*

*This work comes with ABSOLUTELY NO WARRANTY or support.*

***



