---
output:
  html_document:
    keep_md: false
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
params:
  set_title: Blue Whiting Forecasts
---

---
title: `r params$set_title`
---

```{r setup, include="FALSE"}
#'========================================================================
# Visualise skill metrics
#'========================================================================
#
# by Mark R Payne
# DTU-Aqua, Kgs. Lyngby, Denmark
# http://www.staff.dtu.dk/mpay
#
# Created Wed May  1 10:48:18 2019
#
# Makes a series of generic skill visualisations
#
# This work is subject to a Creative Commons "Attribution" "ShareALike" License.
# You are largely free to do what you like with it, so long as you "attribute"
# me for my contribution. See the fine print at the end for exact details.
#
# To do:
#
# Notes:
#   Make sure when knitting that you have the Knit directory set to "project directory"
#'========================================================================

#Load packages
library(PredEng)
library(tidyverse)
library(sf)
library(here)
library(ggnewscale)
#library(scales)

#GGplot config
theme_set(theme_bw(base_size = 14))

#Configure Knitr
knitr::opts_chunk$set(
  echo=FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	results = "markup",
	fig.width = 10,
	fig.height=14/25*10,
  fig.align="center")

#Powerpoint needs 14 x 25cm figures

#knitr::opts_knit$set(root.dir=getwd())
#cat(sprintf("title: %s skill visualisation\n",pcfg@project.name))

```

by Mark R Payne  <br>
DTU-Aqua, Kgs. Lyngby, Denmark  <br>
http://www.staff.dtu.dk/mpay  <br>
```{r results="asis"}
cat(sprintf("Visualisations generated  %s\n",date()))
```


## Configuration

```{r}
#==========================================================================
# Configure
#==========================================================================
pcfg <- readRDS(here("scratch/Blue_whiting_decadal/configuration.rds"))
pcfg <- readRDS(here("scratch/Blue_whiting_NorCPM_tests/configuration.rds"))
show(pcfg)

#'========================================================================
# Setup ####
#'========================================================================
#Import data
this.db <- PE.db.connection(pcfg,results.db = TRUE)
stats.tbl <- tbl(this.db,PE.cfg$db$stats)


```



```{r results="asis"}
#'========================================================================
# Stat skills ####
#'========================================================================
this.sp <- "NorthernComponent"
this.calib <- "MeanAdj"
this.stat <- "SDM"
this.result <- "areaLarvae"
this.stat <- "Mean-salinity"
this.result <- "average"



#Extract the corresponding stat data observations 
stat.obs <- 
  stats.tbl %>%
  filter(srcType=="Observations",
         spName==this.sp,
         statName==this.stat,
         calibrationMethod==this.calib,
         resultName==this.result) %>% 
  collect() %>%
  mutate(year=year(date)) %>%
  filter(month(date) %in% pcfg@MOI) %>%
  mutate(timeframe="Historical")

#Extract the most recent starts in each model
stat.pred <- 
  stats.tbl %>%
  filter(srcType %in% c("Decadal","NMME"),
         spName==this.sp,
         statName==this.stat,
         calibrationMethod==this.calib,
         resultName==this.result,
         !realization %in% c("realmean","ensmean")) %>%
  collect() %>% #Retrieve all
  group_by(srcType,srcName) %>%
  filter(startDate==max(startDate)) %>% #Select the most recent starts
  mutate(type.name=paste(srcType,srcName,sep=".")) %>%
  #Extract bands
  group_by(srcType,srcName,date) %>%
  summarise(tibble(qtile=c("LQ","Med","UQ"),value=quantile(value,probs=c(0.25,0.5,0.75))),
            .groups="drop") %>%
  pivot_wider(names_from="qtile",values_from="value") %>%
  mutate(timeframe="Near-term predictions",
         year=year(date))
      
#Plot timeseries
#g <- 
ggplot(mapping=aes(x=year))+
  geom_ribbon(data=stat.pred,mapping=aes(ymin=LQ,ymax=UQ,fill=srcName),alpha=0.5)+
  geom_line(data=stat.pred,mapping=aes(y=Med,colour=srcName),show.legend = FALSE)+
  labs(fill="Forecast\nmodels")+
  new_scale_colour()+
  geom_line(data=stat.obs,aes(y=value,colour=srcName))+
  labs(x="Year",y="Mean salinity (psu)",
       colour="Historical",
       fill="Forecast\nmodels")

```


## Comparison of lead 1 forecasts

```{r}
#Extract the most recent starts in each model
stat.lead1 <- 
  stats.tbl %>%
  filter(srcType %in% c("Decadal","NMME"),
         spName==this.sp,
         statName==this.stat,
         calibrationMethod==this.calib,
         resultName==this.result,
         realization %in% c("realmean")) %>%
  collect() %>% #Retrieve all
  group_by(srcType,srcName,startDate) %>%
  filter(date==min(date)) %>% #Select the first forecast
  mutate(type.name=paste(srcType,srcName,sep=".")) %>%
  mutate(timeframe="Near-term predictions",
         year=year(date)) %>%
  ungroup()
      
#Plot timeseries
#g <- 
ggplot(mapping=aes(x=year))+
  geom_line(data=stat.lead1,mapping=aes(y=value,colour=srcName,linetype=srcName))+
  labs(fill="Forecast\nmodels")+
  geom_line(data=stat.obs,aes(y=value,colour=srcName,linetype=srcName))+
  labs(x="Year",y="Mean salinity (psu)",
       colour="Data Source",
       linetype="Data Source",
       fill="Forecast\nmodels")

#Check for biases
left_join(y=select(stat.obs,year,value),
          x=select(stat.lead1,year,value),
          by="year") %>%
  mutate(label=sprintf("%02i",year %%100)) %>%
  ggplot(aes(value.x,value.y))+
  geom_abline(slope=1,intercept=0,col="red")+
  geom_path(col="grey90")+
  geom_text(aes(label=label))+
  coord_equal()+
  labs(y="EN4 Observations",
       x="NorCPM realisation mean, lead 3 months")
  


```





***
*This work by Mark R Payne is licensed under a  Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported License. 
For details, see http://creativecommons.org/licenses/by-nc-sa/3.0/deed.en_US
Basically, this means that you are free to "share" and "remix" for 
non-commerical purposes as you see fit, so long as you "attribute" me for my
contribution. Derivatives can be distributed under the same or 
similar license.*

*This work comes with ABSOLUTELY NO WARRANTY or support.*

***



