---
title: "Blue-whiting-salinity-comparisons"
output:
  html_document: 
    df_print: paged
    keep_md: no
    toc: no
    toc_float: yes
---

```{r setup, include="FALSE"}
#'========================================================================
# Blue-whiting-salinity-comparisons
#'========================================================================
#
# by Mark R Payne
# DTU-Aqua, Kgs. Lyngby, Denmark
# http://www.staff.dtu.dk/mpay
#
# Created Mon Mar 22 09:32:17 2021
#
# <Description>
#
# This work is subject to a Creative Commons "Attribution" "ShareALike" License.
# You are largely free to do what you like with it, so long as you "attribute"
# me for my contribution. See the fine print at the end for exact details.
#
# To do:
#
# Notes:
#   * Knitting of notebooks from Rstudio leads to all sorts of messy problems 
#     with the directory. Use here::here() to specify a location relative to
#     the project top directory
#
#'========================================================================

#Configure Knitr
knitr::opts_chunk$set(
  echo=FALSE,
	error = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE,
	results = "markup",
	fig.width = 10,
	fig.height=14/25*10,
  fig.align="center",
  fig.path=here::here("notebooks","figures","Blue-whiting-salinity-comparisons_"))

#Powerpoint needs 14 x 25cm figures

# knitr::opts_knit$set(
#    root.dir="..")

library(tidyverse)
library(ggplot2)
library(tibble)
library(here)
library(RSQLite)
library(lubridate)

theme_set(theme_bw(base_size = 14))
```

```{r}
#==========================================================================
# Configure ####
#==========================================================================

```

by Mark R Payne  <br>
DTU-Aqua, Kgs. Lyngby, Denmark  <br>
http://www.staff.dtu.dk/mpay  <br>
`r format(Sys.time(), "%a %b %d %Y, %H:%M:%S ")`

This notebook makes a direct comparison of the salinity time series extracted from the different data sources. In particular, I suspct that the difference in initialisation data sources may be driving a lot of variability prior to 1980. We look at the first forecasts from the models (i.e. the first March) as a way to get an idea about what is going on with the initialisation of the models.


```{r}
#'========================================================================
# Setup ####
#'========================================================================
db.fname <- here("scratch/Blue-whiting-decadal/Blue-whiting-decadal_Statistics.sqlite")
this.db <- dbConnect(SQLite(),db.fname)
this.tbl <- tbl(this.db,"Statistics")


#Import relevant data
dec.dat <- 
  this.tbl %>%
  filter(srcType=="Decadal",
         !(srcName %in% c("ensmean","Grandens")),
         realization=="realmean",
         lead==2,
         statName=="Mean-salinity") %>% 
  collect()

obs.dat <- 
  this.tbl %>%
  filter(srcType=="Observations",
         statName=="Mean-salinity",
         realization=="realmean",
         calibrationMethod=="MeanAdj") %>%
  collect()

plt.dat <- 
  bind_rows(obs.dat,dec.dat) %>%
  mutate(date=ymd(date)) %>%
  filter(month(date)==3,
         resultName=="average") 


ggplot(plt.dat,aes(x=date,y=value,colour=srcName)) +
  geom_line()+
  facet_wrap(~srcType,ncol=1)+
  labs(x="Year",y="Mean salinity",colour="Product")

```
So, we can see straight away that there is divergence prior to 1980 in both data sets. Plotting this as the sd should make it clearer.

```{r}
plt.dat %>%
  group_by(date) %>%
  summarise(sd=sd(value)) %>%
  ggplot(aes(x=date,y=sd))+
  stat_smooth()+
  geom_point()+
  expand_limits(y=0)+
  labs(x="Year",y="Standard deviation")

```

Which is pretty similar to what we have seen previously. Splitting it up by data sources should confirm this a bit better.

```{r}
plt.dat %>%
  group_by(date,srcType) %>%
  summarise(sd=sd(value)) %>%
  ggplot(aes(x=date,y=sd,colour=srcType))+
  stat_smooth()+
  geom_point()+
  expand_limits(y=0)+
  labs(x="Year",y="Standard deviation")

```

So, yes, that looks more or less right. The increase in the spread of the decadal models after 2000 is odd though - I wonder if this is the same effect again..... But the pre/post 1980 effect is clearly there in both data types, anyway.


***
*This work by Mark R Payne is licensed under a  Creative Commons
Attribution-NonCommercial-ShareAlike 3.0 Unported License. 
For details, see http://creativecommons.org/licenses/by-nc-sa/3.0/deed.en_US
Basically, this means that you are free to "share" and "remix" for 
non-commerical purposes as you see fit, so long as you "attribute" me for my
contribution. Derivatives can be distributed under the same or 
similar license.*

*This work comes with ABSOLUTELY NO WARRANTY or support.*

***



